---
title: "Accesing and Utilizing COVID Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accesing and Utilizing COVID Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction
This vignette provides an overview of how to access and use both COVID case data as well as stringency data. We'll be using two functions from `COVIDutilities` to access both the case counts data from a [JHU GitHub Repo](https://github.com/CSSEGISandData/COVID-19/) as well utilizing an API from [Oxford Blavatnik School of Government](https://covidtrackerapi.bsg.ox.ac.uk).

Let's get started by loading the packages we will need.

```{r setup, message=FALSE,  warning=FALSE}
library(tidyverse)
library(COVIDutilities)
library(glitr)
library(extrafont)
library(scales)
library(lubridate)
```

Pulling both sets of data are pretty easy - you just have to run two lines of code. By default, both `pull_jhu_covid()` and `pull_stringency_index()` are only returning PEPFAR countries. With `pull_stringency_index()`, you have the option of dates to run the API over. We can leave the defaults the same, starting the API from the start of collection (2020-01-22) through today.

```{r import}

  #COVID cases (JHU)
    df_covid <- pull_jhu_covid()

  #Government Response
    df_stringency <- pull_stringency_index()
```

All the functions in `COVIDutilities` convert the names of the countries to the PEPFAR naming conventions so the data can be easily combined with one another as well as with PEPFAR data.

```{r join}
#join datasets
df_viz <- left_join(df_covid, df_stringency)
```

Now that we have the data joined together, we'll plot these together, combining the COVID case counts as an area plot and add a run of the stringency index below that for a few countries.

```{r viz}
 #limit to just South Africa
  df_viz <- df_viz %>%
    filter(countryname == "South Africa",
           date >= "2020-03-01")
df_viz %>% 
    ggplot(aes(date, daily_cases)) +
    geom_col(fill = burnt_sienna, alpha = .8, na.rm = TRUE) +
    geom_hline(aes(yintercept = 0), size = 0.5, color = grey20k) +
    geom_line(aes(y = rollingavg_7day), color = si_palettes$burnt_siennas[7], #size = 1,
              na.rm = TRUE) +
    # geom_vline(xintercept = qtrs, size = 0.5, color = grey20k) +
    geom_rug(aes(color = color), sides="b", na.rm = TRUE) +
    scale_y_continuous(label = comma) +
    scale_x_date(date_labels = "%b %Y",
                 breaks = c(as.Date("2020-03-01"), today())) +
    scale_color_identity() +
    labs(x = NULL, y = NULL, fill = "Stringency Index",
         title = "SOUTH AFRICA CURRENTLY COMING OFF IT'S THIRD WAVE OF COVID-19",
         caption = "Source: Source: JHU COVID-19 feed +  stringency index from Blavatnik School of Government at Oxford University") +
    si_style_nolines()
```

